Modeling of Combined Lead Fast Reactor and Concentrating Solar Power Supercritical Carbon Dioxide Cycles to Demonstrate Feasibility, Efficiency Gains, and Cost Reductions

### 2021-11-10 papers/sustainability-1425053/sustainability-1425053-english.tex
Add files via upload
Make Brian's final changes. Comment out the unneeded statements at the end of the tex on consent statements

% Authors, for the paper (add full first names)
%\dagger -> adds dagger
%\ddagger -> adds double dagger
\Author{\hl{Brian White} %Please carefully check the accuracy of names and affiliations. Changes will not be possible after proofreading.
 $^{1}$, Michael Wagner~$^{1}$\orcidB{}, Ty Neises~$^{3}$, Cory Stansbury~$^{4}$ and Ben Lindley~$^{2,}$*\orcidA{}}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Brian White, Michael Wagner, Ty Neises, Cory Stansbury and Ben Lindley}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{White, B.; Wagner, M.; Neises, T.; Stansbury, C.; Lindley, B.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

@@

\subsection{Cycle Component~Modeling}
%======================================================================================
Components present in the cycles are modeled using various techniques and are discussed in more detail below. Turbines and compressors are analyzed using isentropic efficiencies. Counter-flow heat exchangers (HX) are modeled using the effectiveness-NTU method while simplified ``black box'' heat exchangers that use a simple energy balance for state point calculations are used in lieu of more detailed component models where data are available. The~lead-cooled fast reactor is assumed to be a black box heat exchanger because of the constant heat input and state points on the sCO$_2$ inlet and outlet are provided. The~molten salt loop for the CSP is modeled with necessary components including hot and cold TES, receiver, pumps and counter-flow heat~exchangers.

@@

All of the prior equations are carried out in a built-in function within EES. EES is an iterative solver, and therefore as long as there is a feasible solution, the~functions can take any of the four state points around the heat exchanger and converge on a~solution.

@@

The values displayed in Table~\ref{tab-cycle-constants} are representative of CSP and the Westinghouse LFR. 
%\replaced[comment={Elaborated on table variable names [2,5]}]
{In Table~\ref{tab-cycle-constants}, variable names require further explanation. The~high side pressure with the variable label $P_{2A}$, is the constant pressure outlet on the compressors and inlet to the turbines. In~all models, the value is set by the outlet of the main compressor and held constant with the assumption that there is no pressure drop across heat exchangers. In~addition to this pressure, temperatures are also set. The~inlet temperature of the main compressor, $T_{1A}$, is set to a value of $40~^{\circ}$C in all models. The~temperature of the sCO$_2$ on the outlet of the LFR HX has different variable names, T$_5$, T$_{2C}$, T$_{6A}$, and~T$_{5C}$, according to the associated cycle configuration diagram.}{}.

@@

% \end{paracol}
\begin{figure}[H] 
    %[!h] 
   % \widefigure
    \includegraphics[width=\linewidth]{Definitions/c-lfr-on.pdf}
    \caption{Diagram for C-LFR-ON with focus on electricity~generation\label{c-lfr-on}}
\end{figure}
% \begin{paracol}{2}
% %\linenumbers
% \switchcolumn

@@

The calculated values reported in Table~\ref{tab-noncharg} are component power, mass flow fractions, and~counter-flow heat exchanger specifications. Columns labeled with '$C$' headers have the LFR sCO$_2$ low inlet temperature constrained to 400$~^{\circ}$C while those labeled with '$U$' are unconstrained and calculated based off of a parametric study on the MC mass flow fraction. Values listed in the Cold TES Temperature row are set to values of 390$~^{\circ}$C, 410$~^{\circ}$C, or~440$~^{\circ}$C to study the effects that cold TES temperature has on cycle efficiency and to accommodate certain characteristics of cycle configurations. %\replaced[comment={Removed the conjunction 'and' in the sentence for clarity. [1,1]}]
{Cycles that do not contain the listed component omit the associated values.}
%{Cycles that do not contain the listed component and omit the associated values.} 
\clearpage 
% \mw{Check that the column breakdown and grouping is correct in what I've changed.}
\end{paracol}
\begin{specialtable}[H] 
    %[htbp]

@@

There is an interaction between the temperatures selected here and the non-charging cycle configuration. For~C-LFR-CIRC, the~highest efficiencies are achieved when the cold TES temperature is $10~^{\circ}$C less than the inlet to the LFR, this is due to the approach temperature on the C2S heat exchanger. If~the sCO$_2$ LFR inlet temperature is constrained to $400~^{\circ}$C, then the desired cold TES temperature is $390~^{\circ}$C to allow for the apporach temperature. If~the cold TES is at a higher temperature, then the chiller is needed. This then implies that the sCO$_2$ inlet temperature in the salt-to-sCO$_2$ heat exchanger is  $380~^{\circ}$C. From~the results of the previous section, this is clearly achievable for the C-2HTR3T-ON configuration and the separate cycle configuration, but~for the C-1HTR1T-ON configuration the LFR and CSP inlet temperatures are constrained to be the same. In~this case, the~LFR and CSP inlet temperatures would then be $400~^{\circ}$C, implying a cold TES temperature of  $410~^{\circ}$C and hence a 90\% heat storage efficiency with a requirement for a~chiller.

@@

C-LFR-PAR and C-LFR-CIRC are therefore options going forward, with~a trade-off between heat storage efficiency and component costs, with~the performance of C-LFR-CIRC also depending on the choise of charging~configuration.

@@

 \institutionalreview{\hl{x}}%In this section, please add the Institutional Review Board Statement and approval number for studies involving humans or animals. Please note that the Editorial Office might ask you for further information. Please add ``The study was conducted according to the guidelines of the Declaration of Helsinki, and~approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).'' OR ``Ethical review and approval were waived for this study, due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans or animals. You might also choose to exclude this statement if the study did not involve humans or~animals.

 \informedconsent{\hl{x}}%Any research article describing a study involving humans should contain this statement. Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans. You might also choose to exclude this statement if the study did not involve~humans.

>>>>

% Authors, for the paper (add full first names)
%\dagger -> adds dagger
%\ddagger -> adds double dagger
\Author{\hl{Brian T. White} %Please carefully check the accuracy of names and affiliations. Changes will not be possible after proofreading.
 $^{1}$, Michael J. Wagner~$^{1}$\orcidB{}, Ty Neises~$^{3}$, Cory Stansbury~$^{4}$ and Ben Lindley~$^{2,}$*\orcidA{}}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Brian T. White, Michael J. Wagner, Ty Neises, Cory Stansbury and Ben Lindley}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{White, B.T.; Wagner, M.J.; Neises, T.; Stansbury, C.; Lindley, B.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

@@

\subsection{Cycle Component~Modeling}
%======================================================================================
Components present in the cycles are modeled using various techniques and are discussed in more detail below. Turbines and compressors are analyzed using isentropic efficiencies. Counter-flow heat exchangers (HX) are modeled using the effectiveness-NTU method while simplified ``black box'' heat exchangers that use a simple energy balance for state point calculations are used in lieu of more detailed component models where data are available. The~lead-cooled fast reactor is assumed to be a black box heat exchanger because of the constant heat input and state points on the sCO$_2$ inlet and outlet are provided. The~molten salt loop for the CSP is modeled with necessary components including hot and cold TES, receiver, pumps, and counter-flow heat~exchangers.

@@

All of the prior equations are carried out in a built-in function within EES. EES is an iterative solver, and therefore if there is a feasible solution, the~functions can take any of the four state points around the heat exchanger and converge on a~solution.

@@

The values displayed in Table~\ref{tab-cycle-constants} are representative of CSP and the Westinghouse LFR. 
%\replaced[comment={Elaborated on table variable names [2,5]}]
{In Table~\ref{tab-cycle-constants}, variable names require further explanation. The~high side pressure with the variable label $P_{2A}$, is the constant pressure outlet on the compressors and inlet to the turbines. In~all models, the value is set by the outlet of the main compressor and held constant with the assumption that there is no pressure drop across heat exchangers. In~addition to this pressure, temperatures are also set. The~inlet temperature of the main compressor, $T_{1A}$, is set to a value of $40~^{\circ}$C in all models. The~temperature of the sCO$_2$ on the outlet of the LFR HX has different variable names, T$_5$, T$_{2C}$, T$_{6A}$, and~T$_{5C}$, according to the associated cycle configuration diagram.}{}

@@

% \end{paracol}
\begin{figure}[H] 
    %[!h] 
   \widefigure
    \includegraphics[width=\linewidth]{Definitions/c-lfr-on.pdf}
    \caption{Diagram for C-LFR-ON with focus on electricity~generation\label{c-lfr-on}}
\end{figure}
% \begin{paracol}{2}
% %\linenumbers
% \switchcolumn

@@

The calculated values reported in Table~\ref{tab-noncharg} are component power, mass flow fractions, and~counter-flow heat exchanger specifications. Columns labeled with '$C$' headers have the LFR sCO$_2$ low inlet temperature constrained to 400$~^{\circ}$C while those labeled with '$U$' are unconstrained and calculated based off a parametric study on the MC mass flow fraction. Values listed in the Cold TES Temperature row are set to values of 390$~^{\circ}$C, 410$~^{\circ}$C, or~440$~^{\circ}$C to study the effects that cold TES temperature has on cycle efficiency and to accommodate certain characteristics of cycle configurations. %\replaced[comment={Removed the conjunction 'and' in the sentence for clarity. [1,1]}]
{Cycles that do not contain the listed component omit the associated values.}
%{Cycles that do not contain the listed component and omit the associated values.} 
\clearpage 
% \mw{Check that the column breakdown and grouping is correct in what I've changed.}
\end{paracol}
\begin{specialtable}[H] 
    %[htbp]

@@

There is an interaction between the temperatures selected here and the non-charging cycle configuration. For~C-LFR-CIRC, the~highest efficiencies are achieved when the cold TES temperature is $10~^{\circ}$C less than the inlet to the LFR, this is due to the approach temperature on the C2S heat exchanger. If~the sCO$_2$ LFR inlet temperature is constrained to $400~^{\circ}$C, then the desired cold TES temperature is $390~^{\circ}$C to allow for the approach temperature. If~the cold TES is at a higher temperature, then the chiller is needed. This then implies that the sCO$_2$ inlet temperature in the salt-to-sCO$_2$ heat exchanger is  $380~^{\circ}$C. From~the results of the previous section, this is clearly achievable for the C-2HTR3T-ON configuration and the separate cycle configuration, but~for the C-1HTR1T-ON configuration the LFR and CSP inlet temperatures are constrained to be the same. In~this case, the~LFR and CSP inlet temperatures would then be $400~^{\circ}$C, implying a cold TES temperature of  $410~^{\circ}$C and hence a 90\% heat storage efficiency with a requirement for a~chiller.

@@

C-LFR-PAR and C-LFR-CIRC are therefore options going forward, with~a trade-off between heat storage efficiency and component costs, with~the performance of C-LFR-CIRC also depending on the choice of charging~configuration.

@@

% \institutionalreview{\hl{x}}%In this section, please add the Institutional Review Board Statement and approval number for studies involving humans or animals. Please note that the Editorial Office might ask you for further information. Please add ``The study was conducted according to the guidelines of the Declaration of Helsinki, and~approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).'' OR ``Ethical review and approval were waived for this study, due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans or animals. You might also choose to exclude this statement if the study did not involve humans or~animals.

 %\informedconsent{\hl{x}}%Any research article describing a study involving humans should contain this statement. Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans. You might also choose to exclude this statement if the study did not involve~humans.
